---

---
title: "Lab 5 - Poisson - Questions"
author:
    date: last-modified
    format:
      html:
        self-contained: true
        anchor-sections: true
        code-tools: true
        code-fold: true
        fig-width: 8
        fig-height: 4
        code-block-bg: "#f1f3f5"
        code-block-border-left: "#31BAE9"
        mainfont: Source Sans Pro
        theme: journal
        toc: true
        toc-depth: 3
        toc-location: left
        captions: true
        cap-location: margin
        table-captions: true
        tbl-cap-location: margin
        reference-location: margin
      pdf:
        pdf-engine: lualatex
        toc: false
        number-sections: true
        number-depth: 2
        top-level-division: section
        reference-location: document
        listings: false
        header-includes:
          \usepackage{marginnote, here, relsize, needspace, setspace}
          \def\it{\emph}
execute:
  freeze: auto
  echo: true
  message: false
  warning: false
  fig-align: center
  fig-width: 12
  fig-height: 8
  editor_options: 
  chunk_output_type: inline
  code-overflow: wrap
  html:
    code-fold: true
    code-tools: true
---

1.  To complete this lab:

-   Load packages

```{r}
library(MASS)
library(tidyverse)
library(emmeans)
library(ggeffects)
library(easystats)
library(performance)
library(knitr)
```

-   Download the dataset:

```{r}

library(tidyverse)

data <- read_delim("https://raw.githubusercontent.com/jgeller112/psy504-advanced-stats/main/slides/Poisson/data/2010.csv")

```

2.  Conduct the analysis described in the preregistration document

<!-- -->

a.  The number of hours per week that a person spends on the Internet ("WWWHR") will\
    be predicted by their vocabulary ("WORDSUM"), age ("AGE"), sex ("SEX"), religiosity\
    ("RELITEN"), political orientation ("POLVIEWS"), and how often they work from home\
    ("WRKHOME").

-   Let's use the `naniar` package's function `replace_with_na`to clean the data.

```{r}
library(naniar)

data_pos <- data %>%
  dplyr::select(wwwhr, wordsum, age, sex, reliten, polviews, wrkhome) %>%
replace_with_na(.,
             replace = list(wwwhr = c(-1, 998, 999),
                          wordsum = c(-1, 99),
                          reliten = c(0, 8, 9), 
             polviews = c(0, 8, 9), 
             wrkhome = c(0,8,9), 
             age=c(0, 98, 99)))
```

Q: Can you explain what might be going on in the above code?

**A: It seems like we are using dplyr to replace some codes (like 0, 8, 9, -1, 99, etc) with NAs.**

Q: The next step in data cleaning would be to ensure that the data in your code are aligned with the description/ usage context of the variables

-   Recode sex and reliten as necessary

```{r}
data_pos <- data_pos %>% 
  mutate(sex = recode(sex, `-1` = "Male", `1` = "Female"))

data_pos$sex <- as.factor(data_pos$sex)

data_pos <- data_pos %>% 
    mutate(reliten_recode = factor(reliten, levels = c(1, 2, 3, 4, 5), ordered = TRUE))

```

## Missingness

```{r}
data_pos %>%
  dplyr::select(reliten, reliten_recode)


library(skimr)
skimr::skim(data_pos)

```

## Fit a Poisson model to the data.

a.  The number of hours per week that a person spends on the Internet ("WWWHR") will\
    be predicted by their vocabulary ("WORDSUM"), age ("AGE"), sex ("SEX"), religiosity\
    ("RELITEN"), political orientation ("POLVIEWS"), and how often they work from home\
    ("WRKHOME").

```{r}

model <- glm(wwwhr ~ wordsum + age + sex + reliten_recode + polviews + wrkhome,                  data = data_pos, 
                 family = poisson(link = "log"))

library(broom)
model %>%
  tidy()
```

## Carry out model checking

Hint: performance package has the function you're looking for

```{r}
check_model(model,plot = T)
```

## Find any outliers

```{r}
#Q <- quantile(data_pos$wwwhr, probs=c(.25, .75), na.rm = T)
#iqr <- IQR(data_pos$wwwhr, na.rm = T)

#up <-  Q[2]+1.5*iqr # Upper Range  
#low<- Q[1]-1.5*iqr # Lower Range

#subset(data_pos, data_pos$wwwhr >= (Q[1] - 1.5*iqr) & data_pos$wwwhr <= (Q[2]+1.5*iqr))

# I tried this method above and it cut almost 1000 observations. Any thoughts to why this happened? I'm confused...

stats <- data_pos %>%
  summarize(mean_wwwhr = mean(wwwhr, na.rm = TRUE),
            sd_wwwhr = sd(wwwhr, na.rm = TRUE))

data_pos_filtered <- data_pos %>%
  filter(wwwhr >= (stats$mean_wwwhr - 3 * stats$sd_wwwhr) & 
         wwwhr <= (stats$mean_wwwhr + 3 * stats$sd_wwwhr))



check_outliers(model)
```

## Refit the model after excluding outliers

```{r}
  
  new_model <- glm(wwwhr ~ wordsum + age + sex + reliten_recode + polviews + wrkhome,                  data = data_pos_filtered, 
                 family = poisson(link = "log"))
```

```{r}
model_parameters(new_model) %>%
  print_html()
```

### Check for Overdispersion

Hint: performance package has the function you're looking for

```{r}
check_model(new_model,plot = T)
```

What do you notice? And what's a good next step forward? Can there be another model class that can fit the data? If so, fit this model to the data.

I notice that we no longer have an outlier problem, but we do have an overdispersion problem.

```{r}
check_overdispersion(new_model)
```

## Which one is better- your earlier model, or later model?

```{r}

```

## What is zero inflation? Is there zero-inflation in your chosen model?

```{r}

```

::: panel-tabset
## Log Lambda

```{r}


```

## Mean Count

```{r}

```
:::

## Report your conclusions

We fitted a poisson mixed model

to predict Number_of_fixations with emotion and Group (formula:

Number_of_fixations \~ emotion \* Group). The model's total explanatory power is moderate

(conditional R2 = 0.13) and the part related to the fixed effects alone

(marginal R2) is of 0.02. The model's intercept, corresponding to emotion = 0

and Group = 0, is at 1.35 (95% CI \[1.28, 1.42\], p \< .001). Within this model:

 - The effect of emotion is statistically significant and negative (beta =

-0.09, 95% CI \[-0.16, -0.03\], p = 0.003; Std. beta = -0.04, 95% CI \[-0.08,

-0.01\])

 - The effect of Group is statistically non-significant and positive (beta =

0.09, 95% CI \[-0.05, 0.23\], p = 0.202; Std. beta = 0.04, 95% CI \[-0.02, 0.11\])

 - The effect of emotion × Group is statistically significant and positive (beta

= 0.15, 95% CI \[0.02, 0.27\], p = 0.023; Std. beta = 0.04, 95% CI \[5.19e-03,

0.07\])

Standardized parameters were obtained by fitting the model on a standardized

version of the dataset. 95% Confidence Intervals (CI
